---
title: "p8105_hw3_nk2814_new"
author: "Naama Kipperman"
date: "10/13/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Problem 1 


Load the Instacart data and necessary packages you might need.
```{r}
library(p8105.datasets)
library(tidyverse)
library(dplyr)
data("instacart") 

```
**Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations.**

There are 1,384,617 observations in this instacart dataset and 15 variables - this dataset captures information about online grocery orders placed by Instacart users. Each row in this dataset is a product from a specific order. Examples of variables include order_number, product_name, as well as variables that capture when the order was placed, such as order_dow and order_hour_of_day. For example, a row in this dataset could be an order of Bulgarian Yogurt that was placed at 10 pm on a Saturday, along with more details about that order (the grocery aisle the product came from, whether it was re-ordered, etc.)



**How many aisles are there, and which aisles are the most items ordered from?**

There are 134 distinct aisles.
The aisles from which the most items are ordered are 'fresh vegetables', 'fresh fruits', and 'packaged vegetables fruits'.
