---
title: "p8105_hw3_nk2814_new"
author: "Naama Kipperman"
date: "10/13/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Problem 1 


Load the Instacart data and necessary packages you might need.
```{r}
library(p8105.datasets)
library(tidyverse)
library(dplyr)
data("instacart") 

```
**Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations.**

There are 1,384,617 observations in this instacart dataset and 15 variables - this dataset captures information about online grocery orders placed by Instacart users. Each row in this dataset is a product from a specific order. Examples of variables include order_number, product_name, as well as variables that capture when the order was placed, such as order_dow and order_hour_of_day. For example, a row in this dataset could be an order of Bulgarian Yogurt that was placed at 10 pm on a Saturday, along with more details about that order (the grocery aisle the product came from, whether it was re-ordered, etc.)



**How many aisles are there, and which aisles are the most items ordered from?**

There are 134 distinct aisles.
The aisles from which the most items are ordered are 'fresh vegetables', 'fresh fruits', and 'packaged vegetables fruits'.

```{r}

instacart %>% 
  janitor::clean_names() %>% 
  group_by(aisle) %>%
  summarize(
    number_orders=n()
  ) %>% 
  arrange(desc(number_orders))

```

**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r}

instacart %>% 
  janitor::clean_names() %>% 
  group_by(aisle) %>% 
  summarize(
    number_orders=n()
  ) %>% 
  arrange(desc(number_orders)) %>% 
  filter(number_orders > 10000) %>% 
  ggplot(aes(x=aisle, y=number_orders, fill=aisle))+
  geom_col()+ theme(axis.text.x = element_text(angle = 90))
  

```



**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r}
instacart %>% 
  janitor::clean_names() %>% 
  group_by(aisle, product_name) %>% 
  summarize(
    number_orders=n()) %>% 
  arrange(aisle, desc(number_orders)) %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  top_n(n=3) %>% 
  knitr::kable()
```


**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)**

```{r}

instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_hour_ordered = mean(order_hour_of_day)
  ) %>% 
    mutate(order_dow=recode(order_dow,
       "0"="Sunday",
       "1"="Monday",
       "2"="Tuesday",
       "3"="Wednesday",
       "4"="Thursday",
       "5"="Friday",
       "6"="Saturday")) %>% 
  pivot_wider(
    names_from=order_dow,
    values_from=mean_hour_ordered
  ) %>% 
  knitr::kable()

```




### Problem 2

Load the dataset BRFSS.

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

Do some data cleaning 
- format the data to use appropriate variable names
- focus on the “Overall Health” topic
- include only responses from “Excellent” to “Poor”
- organize responses as a factor taking levels ordered from “Poor” to “Excellent”
```{r}

brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr) %>% 
  separate(locationdesc, into=c("state_abb", "location"), sep=" - ") %>% 
  select(-state_abb) %>% 
  filter(topic == "Overall Health", response %in% c("Poor", "Good", "Fair", "Excellent")) %>% 
  mutate(response = factor(response, order=TRUE)) 


```

In **2002**, the states that were observed at at least 7 locations were  **Connecticut, Florida, Massachusetts, North Carolina, New Jersey, and Pennsylvania**.

In **2010**, the states that were observed at at least 7 locations were **California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, and Washington**. 

```{r}

brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr) %>% 
  separate(locationdesc, into=c("state_abb", "location"), sep=" - ") %>% 
  select(-state_abb) %>%
  group_by(year, state) %>%
  summarize(
    number_locations=n_distinct(location)
  ) %>% 
  filter(number_locations >=7, year %in% c(2002, 2010))
  
  

```

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr) %>% 
  separate(locationdesc, into=c("state_abb", "location"), sep=" - ") %>%
  select(-state_abb) %>% 
  filter(topic == "Overall Health", response =="Excellent") %>% 
  group_by(year, state) %>% 
  mutate(
    avg_data_value = mean(data_value)
  ) %>% 
  select(year, state, avg_data_value) %>% 
  ggplot(aes(x=year, y=avg_data_value, color=state))+
  geom_line(aes(group=state))
  

```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr) %>% 
  separate(locationdesc, into=c("state_abb", "location"), sep=" - ") %>%
  select(-state_abb) %>% 
  filter(topic == "Overall Health", response %in% c("Poor", "Good", "Fair", "Excellent"), year %in% c(2006, 2010), state=="NY") %>% 
    mutate(response = factor(response, order=TRUE)) %>% 
  ggplot(aes(x=location, y=data_value, fill=response)) + geom_bar(position = "dodge", stat="identity") + theme(axis.text.x = element_text(angle = 90)) + facet_grid(year~.)


```



